---
title: "STA303 A1"
author: "Ze Qian Wu (1002192254)"
date: "February 8, 2018"
output:
  word_document: default
  pdf_document: default
---

#Question 1
```{r}
data1 <-read.csv("/Users/Claire/Desktop/bbw.csv")
attach(data1)

maturity=array(0,length(gestation))
MatSmoke=array(0,length(smoke))
for (i in 1:length(gestation))
{
if (gestation[i]<259)
{maturity[i]=1}
else if (gestation[i]>293)
{maturity[i]=3}
else {maturity[i]=2}
}
for (i in 1:length(smoke))
{
if (maturity[i]==1 & smoke[i]==1)
{MatSmoke[i]="PreSmoke"}
else if (maturity[i]==1 & smoke[i]==0)
{MatSmoke[i]="PreNoSmoke"}
else if (maturity[i]==2 & smoke[i]==1)
{MatSmoke[i]="NorSmoke"}
else if (maturity[i]==2 & smoke[i]==0)
{MatSmoke[i]="NorNoSmoke"}
else if (maturity[i]==3 & smoke[i]==1)
{MatSmoke[i]="PostSmoke"}
else {MatSmoke[i]="PostNoSmoke"}
}

v1 = bwt[MatSmoke=="PreSmoke"]
v2 = bwt[MatSmoke=="PreNoSmoke"]
v3 = bwt[MatSmoke=="NorSmoke"]
v4 = bwt[MatSmoke=="NorNoSmoke"]
v5 = bwt[MatSmoke=="PostSmoke"]
v6 = bwt[MatSmoke=="PostNoSmoke"]


boxplot(bwt~smoke, main = "2254")
boxplot(bwt~maturity, main = "2254")
boxplot(v1,v2,v3,v4,v5,v6, names=c("PreSmoke","PreNoSmoke","NorSmoke","NorNoSmoke","PostSmoke","PostNoSmoke"),main = "2254")
```
#In the first graph,there are a few outliers for mothers that did not smoke during pregnacy,however,we can see that the average baby birth weight is lighter if the baby's mother smoked during pregnancy.In the second graph, we can see that the average baby birth weight increases as their maturity level increases. There is a more significant weight difference between level 1 and 2, and there is an outlier lying at level 2.In the third graph, the combined maturity level and maternal smoking status increases when we organize the smoking status in order. PostNoSmoke has the greatest average mean and PreSmoke has the lowest average mean.
#2
```{r}
t.test(bwt~smoke)
```

#By looking at the two sample t-test, the p-value = 4.994e-06, which is significantly smaller than 0.05. Therefore, we reject the p-value and evidence shows that there is a significant difference in the mean birth weight between babies born to mothers who were smokers and babies born to mothers who were nonsmokers.

#Question 3
```{r}
data2 = aov(bwt~factor(maturity))
summary(data2)
TukeyHSD(data2)
#bartlett.test(bwt~maturity)
```
#According to the summary of the one-way analysis of variance, the p-value (<2e-16) is less than 0.05, therefore we reject the p-value and evidence shows that there is a significant difference in the mean birth weight among the levels of maturity. To see which levels of maturity differ,  Tukey's ‘Honest Significant Difference’ method would be appropriate to compare the difference. From the data, we can see that the p-value(3.74e-05) is less than 0.05, which indicates that all maturity levels: 2-1,3-1, and 3-2 have significant difference. 

#Question 4
```{r}
data3 = aov(bwt~factor(MatSmoke))
summary(data3)
TukeyHSD(data3)
```
#According to the summary of the one-way analysis of variance, the p-value (<2e-16) is less than 0.05. Therefore, we reject the p-value and evidence shows that there is a significant difference in mean birth weight among the six categories of babies classified by the combination of their maturity level and mother's smoking status. To see which differ, Tukey's ‘Honest Significant Difference’ method would be appropriate to compare the difference. From the data, any of the pairs that have a p-value less than 0.05 indicates that they have significant difference. 
#(Pairs that have significant difference:NorSmoke-NorNoSmoke, PreNoSmoke-NorNoSmoke, PreSmoke-NorNoSmoke, PostNoSmoke-NorSmoke, PostSmoke-NorSmoke, PreSmoke-NorSmoke, PreNoSmoke-PostNoSmoke, PreSmoke-PostNoSmoke, PreNoSmoke-PostSmoke, PreSmoke-PostSmoke, and PreSmoke-PreNoSmoke)

#Question 5
```{r}
plot(data3, which=2, main="2254")
bartlett.test(bwt~MatSmoke)
```
#Based on the Q-Q plot, the points form a line that is not perfectly straight but very close to the best-fit line, therefore, the graph shows normal distribution. The bartlett test gives a p-value of 0.09627, which is greater than 0.05. In this case, we do not reject the null hypothesis and conclude that all the pairs may have equal variance.


#Question 6
#a) Yes, the number of predictor variables would be the same as in the model used in Q4 which equals to 6. As there are 3 maturity levels and 2 different smoking status, the number of predictor variables would still be 3x2=6
#b) Yes, the F-test for the presence of interaction between maturity level and smoking status would be statistically significant. From the result of Q4, the p-values indicate that all interactions have significant difference.

#Question 7
#According to data we generated, there are no evidence shows that there is a significant difference in the data contained different numbers of babies in the three maturity levels. As ANOVA is robust, we can conclude that we should not concern that the data contained different numbers of babies in the three maturity levels
#Question 8
```{r}
z= lm(bwt ~ smoke + gestation)
summary(z)
```
#The p-value shows that gestation is a significant factor as the null hypothesis is rejected. 
#bwt = a0 + a1 I(smoke==1) + a2 gestation + e(error)
#The difference in models for mean birth weight is that gestation is a numeric but not category variable.

#Question 9
#Two additional potential factors of baby birth weight would be 1) if the mom drinks and 2)if the mom uses marijuana
#The levels would be 1) Never 2)0-10 times before pregnancy 3) 0-10 times during pregnancy 4)more than 10 times before pregnancy 4)more than 10 times during pregnancy
#The levels listed above apply for both factors 1) and 2).
