---

date: "March 21, 2018"
output:
  word_document: default
  html_document: default
---

#part 1
#1
#Geriatric Psychiatry and Neurology
#2
#van der Maas, M., Mann, R. E., McCready, J., Matheson, F. I., Turner, N. E., Hamilton, H. A., . . . Ialomiteanu, A. (2017). Problem gambling in a sample of older adult casino gamblers. Journal of Geriatric Psychiatry and Neurology, 30(1), 3-10. 10.1177/0891988716673468 Retrieved from http://resolver.scholarsportal.info/resolve/08919887/v30i0001/3_pgiasooacg
#3
#Social and Epidemiological Research, Centre for Addiction and Mental Health, University of Toronto, Dalla Lana School of Public Health, 33 Russell Street, Toronto, ON, Canada
#4
# https://journals-scholarsportal-info.myaccess.library.utoronto.ca/details/08919887/v30i0001/3_pgiasooacg.xml
#5
#R 
#6
#Observational, because the assignment mechanism is unknown
#7
#Yes, the article showed a few tables and summary statistics for the collected data and results
#8
#Yes, the article presented normal distribution with p-values of different categories and confidencen intervals of 2.5% and 97.5%.
#9
#2 decimal places
#10
#Logistic Regression Analysis
---------------------------------------------------------------------------------------

#part 2
#1a)
```{r}
videodata <- read.csv("/Users/Claire/Desktop/video.csv")
attach(videodata)
videotable <- table(sex, like)
videotable
chisq.test(videotable, correct = FALSE)
fisher.test(videotable)
```

#According to the summary of both Chi-square test and Fisher's Exact Test, the p-value 0.0680 and 0.0782 are both smaller than 0.1, therefore we reject the p-value and evidence shows that sex is not independent of a student's preference for playing video games. From the conclusion, we can make assumptions that male might have a greater probability of liking video games than women.

#b)
```{r}
x1 <- table(sex[grade == "A"], like[grade == "A"])
x1
chisq.test(x1)
x2 <- table(sex[grade == "nA"], like[grade == "nA"])
x2
chisq.test(x2)
```

#Based on the two chi-square test summary, we can see that the p-value for people getting "A" is smaller than 0.1 whereas the p-value for grade "nA" is greater than 0.1. Therefore, we can conclude that there is evidence showing sex is not independent of a student's preference for playing video games for people getting "A" but it is independent of a student's preference for playing video games for people getting "nA".

#2a)
```{r}
dummysex <- ifelse(sex =="male",1,0)
dummygrade <-ifelse(grade =="A",1,0)
dummylike <- ifelse(like=="yes",1,0)

y1 <- glm(dummylike ~  dummysex + dummygrade + dummysex * dummygrade, family = binomial)
y2 <- glm(dummylike ~  dummysex + dummygrade, family = binomial)

anova(y1,y2, test  = "LRT")
library(aod)
wald.test(Sigma=vcov(y1), b=coef(y1), Terms = 4)
summary(y1)
summary(y2)
```

#The model with the interaction term should be used because the result we obtained from previous questions showed that the variables sex and grade are dependent. To support this assumption, we can use the Wald Test and Likelihood Ratio Test to see if the null hypothesis is valid or not. After conducting the two tests, we can see that they both have a p-value < 0.1, which means that evidence has shown that the null hypothesis is rejected. The null hypothesis = the reduced model, therefore the full model is preferred in this case.  

#b)
#From part a), the data shows that sex and grade have significant interaction effect. The conclusion we draw from the Wald Test and Likelihood Ratio Test showed that the reduced model is rejected, therefore sex is a factor that has an impact on people's grade. Considering the full model is the more appropriate model in this case,it agrees with the answer in 1.b) because it shows that the two variables are not independent.

#3.a)
```{r}
count= c(5,7,1,7,4,22,21,23)
like= c("no","no","no","no","yes","yes","yes","yes")
like= as.factor(like)
sex= c("female","female","male","male","female","female","male","male")
sex = as.factor(sex)
grade= c("A","nA","A","nA","A","nA","A","nA")
grade= as.factor(grade)
gcl=cbind(count,like,sex, grade)
gcl

model3way= glm(count~sex+grade+like+sex*grade+sex*like+grade*like+sex*grade*like, family = poisson)
summary(model3way)
model3way2 = glm(count~sex+grade+like+sex*grade+sex*like+grade*like, family = poisson)
summary(model3way2)

```

#b)
#i
```{r}
anova(model3way, model3way2, test="LRT")
anova(y2, y1, test="LRT")
```
#By using the anova summary table, the p-value for model2.1 and model2.2 is the same as the one for model 3.1 and model 3.2. They both have an indicated p-value of 0.0087, which is smaller than 0.1. Therefore, we reject the null hypothesis of the reduced model and the full model is the preferred model in both cases.

#ii
```{r}
summary(model3way)
summary(y1)

```
#From the summary table, we can use the z-value and its indicated p-value to check if null hypothesis is rejected or not. The two summary tables both have an indicated p-value of 0.01848, which is smaller than 0.1. Therefore, we reject the null hypothesis of the reduced model and the full model is the preferred model in both cases.

#iii
#From the data we obtained from parti)and ii), both Poisson and Logistic Models gave us the same result, which the p-values are all smaller than 0.1 and therefore the null hypothesis of Ho =reduced model is rejected. The deviance gives us a more significant p-value(0.0087) comparing to the Wald test(0.01848), which shows that deviance result is more statistically significant.
